---
title: "Basic Cleaning"
output: html_notebook
---
```{r}
# Load library
library(tidyverse)

# Import original data file
data <- read_csv("university.csv")

```
#BASIC CLEANING
```{r}

# Duplicate rows
sum(duplicated(data)) # None

# Duplicate columns
sum(duplicated(t(data))) # None

# Empty rows
nrow(data[apply(is.na(data) | data=="",1,all),]) # None

# Empty columns
ncol(data %>% select(where(function(x) all(is.na(x) | x=="")))) # None

# Columns with single variable
sum(apply(data,2,function(x) min(x)==max(x))) # None

# NA and blank
sum(is.na(data)) # No NAs in dataset
sum(!grepl("", data)) # No empty strings either.

```

# ALTER VARIABLE NAMES TO STANDARD FORMAT
```{r}
colnames(data) <- str_squish(colnames(data))

# Change spaces to underscores
colnames(data) <-str_replace_all(colnames(data)," ","_")

# Remove trailing underscore
colnames(data) <- str_replace(colnames(data), "_$","")

# Remove 's from Father's and Mother's
colnames(data) <- str_replace(colnames(data),"r's","r")

# Correct spelling of 'Nacionality.
data <- data %>% rename("Nationality" = "Nacionality")

colnames(data) <- tolower(colnames(data))

# Remove / from daytime/evening_attendance
data <- data %>% rename("attendance_mode" = "daytime/evening_attendance")

# Remove brackets
colnames(data) <- str_replace(colnames(data), "\\(", "")
colnames(data) <- str_replace(colnames(data), "\\)", "")

```
# The dataset has three options for outcome, Dropout, Graduate, (still) Enrolled. To simplify the model, I'm going to change this to two options: Dropout or Not Dropout, dropping the Enrolled class.

```{r}
#ADJUST TARGET VARIABLE

# Rename the original 'target' variable as 'outcome'
data <- data %>% rename("outcome" = "target")

# Drop all the rows with outcome 'Enrolled'
data <- subset(data, outcome != "Enrolled")

# Create binary variable 'Dropout'
data <- data %>% mutate (
  dropout = ifelse(outcome == "Dropout",1,0)
)

# Check again that no columns with single variable
sum(apply(data,2,function(x) min(x)==max(x))) #None

```
# NON-RELEVANT COLUMNS

#For this model, I'm only considering pre-enrollment factors (i.e. only information available before enrollment), so all the curricular_unit columns and tuition fees will be deleted. Also the macro-economics factors.
```{r}
data <- data %>% select(-c(
  curricular_units_1st_sem_credited, 
  curricular_units_1st_sem_enrolled, 
  curricular_units_1st_sem_evaluations, 
  curricular_units_1st_sem_approved, 
  curricular_units_1st_sem_grade, 
  curricular_units_1st_sem_without_evaluations, 
  curricular_units_2nd_sem_credited, 
  curricular_units_2nd_sem_enrolled, 
  curricular_units_2nd_sem_evaluations, 
  curricular_units_2nd_sem_approved, 
  curricular_units_2nd_sem_grade, 
  curricular_units_2nd_sem_without_evaluations,
  tuition_fees_up_to_date,
  unemployment_rate,
  inflation_rate,
  gdp,
  application_order
))

```

#Remaining size of the dataset.
```{r}
dim(data)
```


# Export Clean File
```{r}

write.csv(data, "university_clean.csv")

```

